Observation of S-wave scattering of fermions accross a Feshbach resonance

### 2014-12-30 draft1.tex
Draft to send to Ian

\section{Introduction}

\par In order to optimize the signal to noise for low atom numbers, we increased our imaging time.  However, above saturation absorption imaging at sufficiently long exposure times poses a particular problem - if the atoms have had time to absorb enough photons such that the recoil velocity Doppler shifts them from resonance on the order of the linewidth of the resonance, this recoil-induced detuning is a significant correction to the observed absorption of the probe light. In order to take this correction into account, we simulate the imaging process and use the results to extract the atom number and the scattered fraction from our images.
\par This paper is in two parts. In the first, we study absorption imaging in the presence of significant recoil-induced detuning and show how we use our results to interpret data. In the second, we describe our s-wave scattering experiment and, after correcting our data for recoil-induced detuning, extract a measure of the location of the Feshbach resonance in $^{40}K$.

@@

\caption{Using time dependent $I_f$ values obtained from recoil detuning corrected simulation of on-resonant imaging of $^{40}K$ atoms, this graph shows the optical depths each model would deduce from such images. The true optical depth is given at 1.6.$OD_1$ is the high probe intensity corrected optical depth given by \ref{eq2}. $OD_2$ is the model that includes high probe intensity corrections and imaging time corrections from expanding \ref{eq3},\ref{eq4} to second order in imaging time \cite{LJLthesis}. Note that the value obtained using the second order expansion in time starts to differ significantly from the true value after about 20us. The probe intensity is $0.8\, I_{sat}$. }  

@@

@@

Feshbach resonances are a staple tool in the study of cold Fermi gases. Interatomic interactions in Bose-Einstein condensates are easily detected due to the high density of the cloud. However, the density of Fermi clouds is reduced by a factor of $~10^3$ from that of BECs. This makes it necessary to enhance the scattering cross section to observe any effects of interaction. Feshbach resonances allow one to tune the scattering length by changing the magnetic field. Not ony do they enable detection of interactions, but they make it possible to tune interactions from attractive to repulsive, allowing for the phase transition from the BCS to BEC regime at sufficiently low temperatures. 
\par However, to take advantage of Feshbach resonances as a tool it  is first necessary to have a good undestanding of the relationship between the bias magnetic field and the scattering cross section near a Feshbach resonance for specific species and hyperfine states.  The exact value of the resonance is difficult to calculate analytically and must be determined experimentally. Currently, Feshbach resonances are mapped out by observing loss due to three-body inelastic scattering or by re-thermalization or anisotropic expansion measurements that infer the elastic scattering cross section from collective behavior of the cloud \cite{Regal03,OHara02,Monroe93}. 
\par We collide two clouds of $^{40}K$ atoms in a mixture of $\ket{9/2,-9/2}$ and $\ket{9/2,-7/2}$ hyperfine states and directly image the resulting s-wave scattering halo as a function magnetic field strength. This allows us to directly observe the enhancement in scattering without relying on collective behavior. We calculate the fraction of atoms that scattered during the collision, and from this scattering fraction deduce the resonant magnetic field of the resonance.
\par The techniques developed in this experiment for observing Fermion scattering can be extended to engineering higher order partial wave interactions, as has been done for bosons \cite{Williams2012}.

@@

\bibliography{refs}{}
\bibliographystyle{plain}

>>>>

\section{Introduction}
Feshbach resonances are a staple tool in the study of cold Fermi gases. Interatomic interactions in Bose-Einstein condensates are easily detected due to the high density of the cloud. However, the density of Fermi clouds is reduced by a factor of $~10^3$ from that of BECs. This makes it necessary to enhance the scattering cross section to observe any effects of interaction. Feshbach resonances allow one to tune the scattering length by changing the magnetic field. Not ony do they enable detection of interactions, but they make it possible to tune interactions from attractive to repulsive, allowing for the phase transition from the BCS to BEC regime at sufficiently low temperatures. 
\par However, to take advantage of Feshbach resonances as a tool it  is first necessary to have a good undestanding of the relationship between the bias magnetic field and the scattering cross section near a Feshbach resonance for specific species and hyperfine states.  The exact value of the resonance is difficult to calculate analytically and must be determined experimentally. Currently, Feshbach resonances are mapped out by observing loss due to three-body inelastic scattering or by re-thermalization or anisotropic expansion measurements that infer the elastic scattering cross section from collective behavior of the cloud \cite{Regal03,OHara02,Monroe93}. 
\par We collide two clouds of $^{40}K$ atoms in a mixture of $\ket{9/2,-9/2}$ and $\ket{9/2,-7/2}$ hyperfine states and directly image the resulting s-wave scattering halo as a function magnetic field strength. This allows us to directly observe the enhancement in scattering without relying on collective behavior. We calculate the fraction of atoms that scattered during the collision, and from this scattering fraction deduce the resonant magnetic field  and width of the resonance.
\par The techniques developed in this experiment for observing Fermion scattering can be extended to engineering higher order partial wave interactions, as has been done for bosons \cite{Williams2012}.
\par However, even with Feshbach resonance enhancement of the scattering cross section, Fermi gases are dilute enough that very few atoms will scatter as a fraction of total atom number. This makes direct detection of s-wave scattering halo difficult due to detection uncertainty, which dispraportionally effects low atom numbers. In order to optimize the signal to noise for low atom numbers, we increased our imaging time.  However, above saturation absorption imaging at sufficiently long exposure times poses a particular problem - if the atoms have had time to absorb enough photons such that the recoil velocity Doppler shifts them from resonance on the order of the linewidth of the resonance, this recoil-induced detuning is a significant correction to the observed absorption of the probe light. In order to take this correction into account, we simulate the imaging process and use the results to extract the atom number and the scattered fraction from our images.
\par This paper is in two parts. In the first, we study absorption imaging in the presence of significant recoil-induced detuning and show how we use our results to interpret data. In the second, we describe our s-wave scattering experiment and, after correcting our data for recoil-induced detuning, extract a measure of the location of the Feshbach resonance in $^{40}K$.

@@

\caption{Using time dependent $I_f$ values obtained from recoil detuning corrected simulation of on-resonant imaging of $^{40}K$ atoms, this graph shows the optical depths each model would deduce from such images. The true optical depth is given at 1.6.$OD_1$ is the high probe intensity corrected optical depth given by \ref{eq2}. $OD_2$ is the model that includes high probe intensity corrections and imaging time corrections from expanding \ref{eq3},\ref{eq4} to second order in imaging time \cite{LJLthesis}. Note that the value obtained using the second order expansion in time starts to differ significantly from the true value after about 15us. The probe intensity is $0.8\, I_{sat}$. }  

@@

Now that we have the ability to interpret absorption images at larger imaging times, we can choose an imaging time based on optimal signal to noise ratio. Here, we consider only the photon shot noise that results in camera count noise, with the uncertainty proportional to $\sqrt{N_p}$, where $N_p$ is the photon number. We then propagate this uncertainty through our correction scheme to deduce the final uncertainty in the calculated optical depth. 
\par As seen in figure \ref{fig:SNR}, around 40$\mu s$ for reasonable parameters the uncertainty levels out and there are no longer appreciable benefits by extending the imaging time. The advantage of going to a 40$\mu s$ as opposed to 10$\mu s$ where the uncorrected model is appropriate is about a factor of  1.6 reduction in measurement uncertainty. 

@@

In this section we describe our experiment with scattering two counter-propagating clouds of $^{40}K$ atoms and observe the resulting s-wave halo of scattered atoms. We vary the magnetic field around the Feshbach resonance during this experiment, and from the change in scattering deduce a location of the magnetic fields resonance of 202.47 Gauss and a width of 8.98 Gauss.

@@

We studied the effects of recoil-induced detuning effects on absorption images and found an optimal imaging time of ~40$\mu s$ for $^{40}K$ atoms for noise minimization after corrections. We use these results to observe s-wave scattering halos of the Fermi gas around the Feshbach resonance and directly verify the resonance location and width. Our analysis can be used in any absorption imaging applications where signal to noise minimization is critical. Experimental techniques devoloped for our s-wave scattering expeiment can be readily extended to a Raman-coupled single component gas version, in which we hope to detect synthetic p-wave interactions. 
\section*{Acknowledgements}
We thank Marcell Gall for helpful discussions. This work was supported by ???

\section*{References}
\bibliography{refs}{}
\bibliographystyle{plain}
